---
title: "01-exercises"
author: "Jennifer Lee"
date: "March 31, 2017"
output: html_document
---


## Exercise 1:

Write a function (f) that takes a vector of numbers, **x**, and returns a vector of numbers such that each element containing the product of every element of x except the element of x with the same index. 


### Example 

    > x <- c( 1, 5, 2, 8 ) 
    > f(x) 
    [1] 80 16 40 10
    # 5*2*8, 1*2*8, 1*5*8, 1*2*5  

### Solution 

```{r}
x <- c(1, 5, 2, 8)
f <- function(x) {
  result <- 0
  for(i in 1:4) {
    result[i] <- prod(x[-i])
  }
  return(result)
}
# Alternative solution:
# prod(x)/x
#R is a vector language that is why that works
f(x)
```
    


## Exercise 2

Write a function f(x) to accept an integer vector, and returns a vector with those numbers except for: 
   multiples of 3 replaced by "Fizz"? 
   multiples of 5 replaced by "Buzz" 
   multiples of 3 and 5 replaced by "FizzBuzz""
   
### Example

    > x <- 1:20
    > f(1:20) 
    # 1 2 Fizz 4 Buzz Fizz 7 8 Fizz Buzz 11 Fizz 13 14 FizzBuzz 16 17 Fizz 19 Buzz

### Solution

```{r}
x <- 1:20
f <- function(x) {
ifelse(x %% 3 == 0 & x %% 5 == 0, "FizzBuzz",
       ifelse(x %% 3 == 0, "Fizz", 
              ifelse(x %% 5 == 0, "Buzz", x)
              )
       )
  
}
f(1:20)
```
