---
title: "Classification Mettrics"
author: "Jennifer Lee"
date: "May 7, 2017"
output: html_document
---


**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file.** 
 

## Classification Metrics Functions 

Write functions of `y` and `y_hat` to calculate the following classification metrics

* True Positive Rate  
* False Positive Rate
* True Negative Rate  
* False Negative Rate 
* Sensitivity 
* Specificity 
* Recall 
* Precision
* Prevalence 
* Accuracy
* Kappa

```{r "definitions", echo=FALSE}
#Predicted conditions are y_hat
#Observed conditions are y
#True Positive (TP) = y positive and y_hat positive
#False Negative (FN) = y positive and y_hat negative
#False Positive (FP) = y negative and y_hat positive
#True Negative (TN) = y negative and y_hat negative

#True Positive Rate = sensitivity or recall
TPR <- function(TP, FN) {
  TP/(TP+FN) }

#False Positive Rate
FPR <- function(FP, TN) {
  FP/(FP+TN) }

#True Negative Rate = specificity
TNR <- function(TN, FP) {
  TN/(TN+FP) }

#False Negative Rate
FNR <- function(FN, TP) {
  FN/(FN+TP) }

#Sensitivity = true positive rate
sensitivity <- function(TP, FN) {
  TP/(TP+FN) }

#Specificity
specificity <- function(TN, FP) {
  TN/(TN+FP) }

#Recall = sensitivity or true positive rate
recall <- function(TP, FN) {
  TP/(TP+FN) }

#Precision
precision <- function(TP, FP) {
  TP/(TP+FP) }

#Prevalence
prevalence <- function(TP, FN, FP, TN) {
  (TP+FN)/(TP+FN+FP+TN) }

#Accuracy
accuracy <- function(TP,TN,FP,FN) {
  (TP+TN)/(TP+FP+FN+TN) }

#Kappa
kappa <- function(TP,FP,FN,TN) {
  p_0 <- (TP+TN)/(TP+FP+FN+TN)
  p_yes <- ((TP+FP)/(TP+FP+FN+TN))*((TP+FN)/(TP+FP+FN+TN))
  p_no <- ((FN+TN)/(FN+TN+FP+TP))*((TN+FP)/(TP+FP+FN+TN))
  p_e <- p_yes + p_no 
  k <- (p_0-p_e)/(1-p_e) 
  return(k) }
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


## Usage 

Use the function(s) above to calculate the metrics for your late arrival model that
you created last week. 

```{r "evaluations"}
library(readr)
library(dplyr)
#Read the following csv files from data folder from week 2 of class
airports <- read_csv("airports.csv")
flights <- read_csv("flights.csv")
planes <- read_csv("planes.csv")
weather <- read_csv("weather.csv")
#Combine datasets into one data table:
combined <- left_join(flights, planes, by = "tailnum")
combined <- left_join(combined, airports, c("origin" = "faa"))
combined <- left_join(combined, airports, c("dest" = "faa"))
combined <- left_join(combined, weather, by = c("origin" = "origin", "time_hour" = "time_hour"))
#There are 9,430 NA's out of 336,776 rows, delete the NAs
subsetcombined <- combined[complete.cases(combined$arr_delay),]
#Keep only columns with numeric data
subsetcombined <- subset(combined, select = c(month.x, dep_time, dep_delay, arr_time, arr_delay, air_time, distance, hour.x, temp, dewp, humid, wind_dir, wind_speed, wind_gust, precip, pressure, visib))
#Create a new column called "delay22min" with two categories: flights with an arrival delay >= 22 min and flights with arrival delay < 22 min
subsetcombined$delay22min <- ifelse(subsetcombined$arr_delay >= 22, "delay >= 22 min", "delay < 22min") 
subsetcombined$delay22min <- as.factor(subsetcombined$delay22min)
subsetcombined <- subsetcombined[complete.cases(subsetcombined$delay22min),]

# setting seed to reproduce results of random sampling
set.seed(1000)
#row indices for training data
ind <- sample(2, nrow(subsetcombined), replace=TRUE, prob=c(0.7,0.3))
#model training data
traindataset <- subsetcombined[ind==1,]
#test data
testdataset <- subsetcombined[ind==2,]
testdataset$delay22min <- as.factor(testdataset$delay22min)

#build the model
results <- lm(arr_delay~., testdataset)
summary(results)

#prediction
pred <- predict(results, newdata=testdataset, type = "response")

#check results
head(pred)
head(testdataset$arr_delay)
head(testdataset$delay22min)

#label predictions as either Early/On-Time if arrival delay is < 22 min
model_pred_direction <- rep("Early/On-Time", 97564)
model_pred_direction[pred > 22] <- "Late"

table(model_pred_direction, testdataset$delay22min)

TP <- 77747
TN <- 14302
FN <- 444
FP <- 5114

#True Positive Rate = sensitivity or recall
TPR <- function(TP, FN) {
  TP/(TP+FN) }
TPR(TP, FN)

#False Positive Rate
FPR <- function(FP, TN) {
  FP/(FP+TN) }
FPR(FP, TN)

#True Negative Rate = specificity
TNR <- function(TN, FP) {
  TN/(TN+FP) }
TNR(TN,FP)

#False Negative Rate
FNR <- function(FN, TP) {
  FN/(FN+TP) }
FNR(FN, TP)

#Sensitivity = true positive rate
sensitivity <- function(TP, FN) {
  TP/(TP+FN) }
sensitivity(TP, FN)

#Specificity
specificity <- function(TN, FP) {
  TN/(TN+FP) }
specificity(TN, FP)

#Recall = sensitivity or true positive rate
recall <- function(TP, FN) {
  TP/(TP+FN) }
recall(TP, FN)

#Precision
precision <- function(TP, FP) {
  TP/(TP+FP) }
precision(TP, FP)

#Prevalence
prevalence <- function(TP, FN, FP, TN) {
  (TP+FN)/(TP+FN+FP+TN) }
prevalence(TP,FN,FP,TN)

#Accuracy
accuracy <- function(TP,TN,FP,FN) {
  (TP+TN)/(TP+FP+FN+TN) }
accuracy(TP,TN,FP,FN)

#Kappa
kappa <- function(TP,FP,FN,TN) {
  p_0 <- (TP+TN)/(TP+FP+FN+TN)
  p_yes <- ((TP+FP)/(TP+FP+FN+TN))*((TP+FN)/(TP+FP+FN+TN))
  p_no <- ((FN+TN)/(FN+TN+FP+TP))*((TN+FP)/(TP+FP+FN+TN))
  p_e <- p_yes + p_no 
  k <- (p_0-p_e)/(1-p_e) 
  return(k) }
kappa(TP,FP,FN,TN)
```

