---
title: "Assingment NYC Flight Data"
author: "Jennifer Lee"
date: "April 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. 
In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, 
`airports.csv`, `planes.csv` and `weather.csv`.  Join/merge these tables such 
that there is one rectangular table with one row for each flight. 


Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

```{r}
#Load readr
library(readr)

#Set working directory
setwd("~/machine_learning_r/02-building-blocks/02-exercise-nycflights/data")
#NEXT TIME DO NOT SET WORKING DIRECTORY TO OWN FOLDER!!!

#Read the following csv files from data folder
airports <- read_csv("data/airports.csv")
flights <- read_csv("data/flights.csv")
planes <- read_csv("data/planes.csv")
weather <- read_csv("data/weather.csv")
```

## Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}
#Load ggplot2
library(ggplot2)

#Plot histogram of arrival delays
arrdelayhist <- qplot(flights$arr_delay,
      geom="histogram",
      main = "Histogram of Arrival Delays for NYC Flights",
      xlab = "Arrival Delay (minutes)",
      ylab = "Count",
      fill=I("cyan4"),
      binwidth = 30,
      ylim = c(0, 300000))
arrdelayhist

#Plot histogram of departure delays
depdelayhist <- qplot(flights$dep_delay,
      geom="histogram",
      main = "Histogram of Departure Delays for NYC Flights",
      xlab = "Departure Delay (minutes)",
      ylab = "Count",
      fill=I("darkslateblue"),
      binwidth = 100)
depdelayhist
```

## Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}
#Combine year, month, and day columns into one column labeled "date"
flights$date <- as.Date(with(flights, paste(year, month, day,sep="-")), "%Y-%m-%d")

#Scatterplot using qplot for date vs arrival delays
scatter <- qplot(date, arr_delay, data = flights, 
      colour = I("darkorange3"),
      xlab = "Date",
      ylab = "Arrival Delay (minutes)",
      main = "Date vs Arrival Delay for Flights in NYC")
scatter
```



## Create tables for Categorical Variables

Create a `table`, counting the flights by origin airports and carrier.

```{r}
#Load plyr package
library(plyr)
#Count flights by origin and carrier
origincarrier <- count(flights, c('origin', 'carrier'))

#Alternative solution:
library(dplyr)
# create data frame grouped by origin and carrier
g <- group_by(flights.df, origin, carrier)

# use View() to view the entire table
View(origincarrier)

```


## Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}
#Load dplyr
library(dplyr)

#Do a left join with flights and planes data sets and join by variable tailnum
combined <- left_join(flights, planes, by = "tailnum")

#Do another left join with the previous combined data set and airport by variable origin
combined <- left_join(combined, airports, c("origin" = "faa"))

#Repeat above set this time joining by destination, need to separate this because it has to be an OR join not AND
combined <- left_join(combined, airports, c("dest" = "faa"))

#Last left join, combine above with weather data set and join by origin AND date
combined <- left_join(combined, weather, by = c("origin" = "origin", "time_hour" = "time_hour"))

View(combined)

#L <- list(airports, flights, planes, weather)
#join_all(L, type='inner')

#Professor's solution:
flights$carrier %>% table( useNA="always") %>% sort
flights$origin %>% table( useNA="always") %>% sort

yx <- flights
yx <- merge( yx, planes, by="tailnum", all.x = TRUE)
yx <- merge( yx, airports, by.x="origin", by.y="faa", all.x=TRUE, suffixes=c("",".origin") ) #suffix adds a suffix to end of column name, don't want to replace the first part of the column name so leave that argument blank (i.e., "")
yx <- merge(yx, airports, by.x="dest", by.y="faa", all.x=TRUE, suffixes=c("", ".dest"))
yx <- merge(flights, weather, by=c("origin", "year", "month", "day", "hour"), all.x=TRUE, suffixes=c('',".weather"))

```



