---
title: "NYCFlights: Arrival Delay Regression Model (Open Ended Variation)"
author: "Jennifer Lee"
date: "April 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NYCFlights Model

Using the rectangular data that you created last week create a model for `arr_delay`. Follow a well defined process for creating a model as discussed in class or from one of the readings. 

Break the process down into steps and create a codeblock for each step. 
 - Use as many steps you feel as necessary
 - Concisely explain each step 
 - Show all work
 
(You can start with the data that you created last week.)


## Step 1: Load packages and data into R

```{r}
# Load the following packages into R
library(readr)
library(dplyr)
library(caret)

#Read the following csv files from data folder from week 2 of class
airports <- read_csv("airports.csv")
flights <- read_csv("flights.csv")
planes <- read_csv("planes.csv")
weather <- read_csv("weather.csv")
```

## Step 2: Pre-processing Data
```{r}
#Combine datasets into one data table:

#Do a left join with flights and planes data sets and join by variable tailnum
combined <- left_join(flights, planes, by = "tailnum")
#Do another left join with the previous combined data set and airport by variable origin
combined <- left_join(combined, airports, c("origin" = "faa"))
#Repeat above set this time joining by destination, need to separate this because it has to be an OR join not AND
combined <- left_join(combined, airports, c("dest" = "faa"))
#Last left join, combine above with weather data set and join by origin AND date
combined <- left_join(combined, weather, by = c("origin" = "origin", "time_hour" = "time_hour"))

#Check for NA's in arrival delay column
summary(combined$arr_delay)
#There are 9,430 NA's out of 336,776 rows, delete the NAs
subsetcombined <- combined[!is.na(combined$arr_delay),]
#Keep only columns with numeric data
subsetcombined <- subset(subsetcombined, select = c(month.x, dep_time, dep_delay, arr_time, arr_delay, air_time, distance, hour.x, temp, dewp, humid, wind_dir, wind_speed, wind_gust, precip, pressure, visib))

summary(subsetcombined)

```

## Step 3: Create training and test data
```{r}
# setting seed to reproduce results of random sampling
set.seed(100)
#row indices for training data
trainingRowIndex <- sample(1:nrow(subsetcombined), 0.7*nrow(subsetcombined)) 
#model training data
trainingData <- subsetcombined[trainingRowIndex, ] 
#test data
testData <- subsetcombined[-trainingRowIndex, ]
```

## Step 4: Develop model on training data and use to predict on test data
```{r}
#build the model
lmMod <- lm(arr_delay ~ ., data=trainingData)
#predict arrival delay
arrdelayPred <- predict(lmMod, newdata=testData)

#review diagnostic measures
summary(lmMod)
```
The following variables have a low p-value (i.e. at least a 95% confidence interval): month, departure delay, arrival time, air time, distance, hour, temperature, dew point, humidity, wind direction, precipitation, pressure, and visibility,

The R-squared value is 0.8788, which means there is a positive correlation between arrival delay and all the variables.


## Step 5: Calculate prediction accuracy and error rates
```{r}

#create a data frame comparing the observed values from the test data to the predicted values from the predicted values
modelvalues <- data.frame(obs = testData$arr_delay, pred=arrdelayPred)
#show summary results for Root Mean Square Error and R-squared values
defaultSummary(modelvalues)
```

# Question:

Is this a good model?

The root mean square error and the r-squared values for the test and predicted values are 14.75 and 0.878, respectively. Since the r-squared value is fairly high, I would say that the test data is a good model of the training data.
